### 2024-01-23 outputs/paper/paper.qmd
initial final paper upload

---
title: "Toronto's Most Studious Wards:  An Analysis of Annual Toronto Public Library's Circulation and Visits by Wards Prior to and Post-COVID"
subtitle: "Data Visualization in R"
author:
  - Francesca Ye
email: francesca.ye@mail.utoronto.ca
prerequisites: "Datasets taken from opensourcetoronto and can be found here: https://open.toronto.ca/catalogue/"
date: January 22 2024 
date-format: long
abstract: "test"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---
```{r}
library(tinytex)
```

# Introduction

# Data
## Library Branch General Information
## Library Circulation
## Library Visits
## Annual Library Circulation and Visits by Ward
## City-Wide Annual Circulation and Visits

# Results
## Annual Circulation by Ward
## Annual Visits by Ward
## City-Wide Annual Circulation and Visits

# Discussion

# Conclusion